---
title: "NHL Game Predictions"
author: "Marcus Becker"
date: "October 13, 2019"
output: html_document
---

```{r setup, include=FALSE}

# Load packages
library(tidyverse)
library(broom)
library(ranger)
library(mgcv)
library(rsample)
library(fs)
library(lubridate)
library(here)
library(RcppRoll)
library(plotly)


# Import data - also, we want to be able to scrape this data easily and on an ongoing basis. Learn rvest package.

paths <- dir_ls(here("./data/base"), glab = "*.csv")

names <- c("reg_1516_es",
           "reg_1617_es",
           "reg_1718_es",
           "reg_1819_es")

ls_raw <- map(paths, read_csv) %>%
  set_names(names)

```

We begin with the following data:

+ NHL regular-season games for four seasons: 2015-16 through 2018-19. 
+ All data is even-strength, non-score and venue adjusted. 
  - So this includes 5v5, 4v4, and 3v3 play.

Our goal is to build a model capable of predicting future NHL games.

Here are some ideas for predictors:

+ Pts percentage (of both teams)
+ Home vs. Away game
+ Scheduling
  - What was the distance travelled since the last game? Back to back games?
+ Advanced metrics like CF, FF, xGF, etc. 
  - Probably want to build in here some variables that capture how well a team has been playing recently.
+ Injuries? Pct of team cap that is playing in the game. 
  - Data source: capgeek.
+ Special teams
  - PP/PK ... maybe factor in a recency thing as well.
  
Other considerations on how to build this ..

How about constructing a lookup table for each teams performance, by date.

Date | Team | metrics, etc. 

Then build the train/test dataset off of that. 

```{r}

# Let's look at one team in 15-16 to start. Oilers!

oilers_1516 <- ls_raw[["reg_1516_es"]] %>%
  filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA) %>%
  # Create Date variable
  mutate(Date = ymd(str_extract(Game, "^.{10}"))) %>%
  separate(Game, into = c("date", "score"), sep = " - ", remove = FALSE) %>%
  mutate(home = ifelse(str_detect(score, "^Oilers"), 0, 1)) %>%
  mutate(score_numeric = str_match_all(score, "[0-9]+")) %>%
  mutate(score_numeric = map(score_numeric, setNames, c("away_score", "home_score"))) %>%
  unnest_wider(score_numeric) %>%
  # Code game outcome:
  mutate(outcome = as.factor(ifelse((home_score > away_score & home == "1") |
                                    (away_score > home_score & home == "0"), 
                                    "Win", "Loss")))

```

Now we need to scale this to all teams. This is for 2,460 games of the regular season in 15-16.

```{r warning=FALSE}

allteams_1516 <- ls_raw[["reg_1516_es"]] %>%
  # filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA) %>%
  # Create Date variable
  mutate(Date = ymd(str_extract(Game, "^.{10}"))) %>%
  separate(Game, into = c("date", "score"), sep = " - ", remove = FALSE) %>%
  # mutate(home = ifelse(str_detect(score, "^Oilers"), 0, 1)) %>%
  group_by(Game) %>%
  mutate(home = seq(0, 1, 1)) %>%
  mutate(score_numeric = str_match_all(score, "[0-9]+")) %>%
  mutate(score_numeric = map(score_numeric, setNames, c("away_score", "home_score"))) %>%
  unnest_wider(score_numeric) %>%
  # Code game outcome:
  mutate(outcome = ifelse((home_score > away_score & home == "1") | (away_score > home_score & home == "0"), "Win", "Loss"))

```

Now, can we scale to all four season?!

```{r message = FALSE, warning = FALSE}

df_1 <- ls_raw[["reg_1516_es"]] %>%
  # filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA) 
df_2 <- ls_raw[["reg_1617_es"]] %>%
  # filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA) 
df_3 <- ls_raw[["reg_1718_es"]] %>%
  # filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA) 
df_4 <- ls_raw[["reg_1819_es"]] %>%
  # filter(Team == "Edmonton Oilers") %>%
  select(Game, Team, CF, CA, FF, FA, SF, SA, GF, GA, xGF, xGA, SCF, SCA,
         HDCF, HDCA)

df_master <- dplyr::bind_rows(df_1, df_3)
  
str(df_raw)

```


```{r}

# Code the rolling averages of corsi, fenwick, shots, and goals. 

allteams_1516_rollin <- allteams_1516 %>%
  # Select the parameters we're interested in:
  select(Game, Date, Team, CF, CA, FF, FA, SF, SA, GF, GA) %>%
  # Set up 'windows' of games
  crossing(window = c(3, 5, 10, 20)) %>%
  group_by(window, Team) %>%
  nest() %>%
  # Corsi
  mutate(rolling_CF = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$CF, n = .y, fill = NA, align = "right")),
         rolling_CA = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$CA, n = .y, fill = NA, align = "right")),
         rolling_C_pct = map2(.x = rolling_CF, 
                            .y = rolling_CA, 
                            .f = ~ (.x / (.x + .y)))) %>%
  # Fenwick
  mutate(rolling_FF = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$FF, n = .y, fill = NA, align = "right")),
         rolling_FA = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$FA, n = .y, fill = NA, align = "right")),
         rolling_F_pct = map2(.x = rolling_FF, 
                            .y = rolling_FA, 
                            .f = ~ (.x / (.x + .y)))) %>%
  # Shots
  mutate(rolling_SF = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$SF, n = .y, fill = NA, align = "right")),
         rolling_SA = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$SA, n = .y, fill = NA, align = "right")),
         rolling_S_pct = map2(.x = rolling_SF, 
                            .y = rolling_SA, 
                            .f = ~ (.x / (.x + .y)))) %>%
  # Goals
  mutate(rolling_GF = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$GF, n = .y, fill = NA, align = "right")),
         rolling_GA = map2(.x = data, 
                           .y = window, 
                           .f = ~ roll_sum(x = .x$GA, n = .y, fill = NA, align = "right")),
         rolling_G_pct = map2(.x = rolling_GF, 
                            .y = rolling_GA, 
                            .f = ~ (.x / (.x + .y)))) %>%
  # Select the rolling percentages
  select(window, Team, data, rolling_C_pct, rolling_F_pct, rolling_S_pct) %>%
  unnest(c(rolling_C_pct, rolling_F_pct, rolling_S_pct)) %>%
  mutate(game_num = seq(1, 82, 1)) %>%
  ungroup(window, Team)

# Let's view the Canadian Teams graphically!

cdn_teams <- c("Edmonton Oilers", "Vancouver Canucks", "Calgary Flames", "Winnipeg Jets",
               "Toronto Maple Leafs", "Ottawa Senators", "Montreal Canadiens")

plot1 <- allteams_1516_rollin %>%
  mutate(window = as.factor(window)) %>%
  mutate(window = fct_infreq(window)) %>%
  filter(Team %in% cdn_teams) %>%
  # Choosing corsi
  ggplot(aes(x = game_num, y = rolling_C_pct, color = Team)) +
  geom_line() +
  geom_point() +
  geom_hline(aes(yintercept = 0.5), color = "black", lty = 2) +
  facet_wrap(~ window)


ggplotly(plot1)
  




```


















