---
title: "Defensive R Programming"
author: "Marcus Becker"
date: "March 31, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 1: Avoiding Conflict

- Beware when updating packages -> you could fix serious bugs, but also you could potentially be breaking working code.
- Use Task Views (from CRAN)
- When working in R, objects that are created are placed in the global environment
- Packages use *namespaces* as spaces for name - you can think of a namespace as a box that contains the package.
- A namespace helps keep things tidy - sort of like folders.
- Namespace clashes are challenging! e.g. filter.

```{r}

# Return a vector of exported functions:
length(getNamespaceExports("dplyr")) # 262 exported functions!
length(getNamespaceExports("abmi.camera.extras")) # Three exported functions :/ lol.
# What about the tidyverse in total ... ?
length(getNamespaceExports("tidyverse")) # doesn't quick work the same. 

search()

library(conflicted) # this package is now bumped to the top of the search path.

library(dplyr)

filter() # Throws are error! Conflicted forces us to make our call of filter explicit, ie dplyr::filter or stats::filter

```

# Chapter 2: Early Warning Systems

- Boolean operators TRUE/FALSE -> special values in R. Can't be overrided. Watch out for T/F.
- the message() function: signals to the user the state of a process; contains helpful information; very handy for long running processes.

```{r}

suppressPackageStartupMessages(library(dplyr))

# Supressing messages
sim = suppressMessages(simulate(runs = 5))

N <- 10
results <- numeric(N)
for (i in 1:N) {
  results[i] <- simulate()
  # Show a progress report with a message()
  message("Simulation", i, "complete!")
}

# Writing a function
get_distribution <- function(N, verbose = TRUE) {
  results <- numeric(N)
  for(i in 1:N) {
    results[i] <- simulate()
    # Check if verbose = TRUE
    if(isTRUE(verbose)) {
      # Show a progress report
      message("Simulation", i, "complete!")
    }
  }
  return(results)
}

get_distribution(N = 10, verbose = TRUE)

```

- The warning message. Signals that something may have gone wrong, but R continues. 

```{r}

# Create new variable
x <- c(1,1,1)
y <- 1:3

# Calculate correlation coefficient
m <- cor(x,y) # get a warning

suppressWarnings(cor(x,y))

```

- Stop! Sometimes things are broken. We need to raise an error.
- You can't suppress (or ignore) errors. The definition of an error is that R can't continue. Instead, we *catch* errors.

```{r}

# Stop the execution if ages are negative
mean_age <- function(ages) {
  if(any(ages < 0 | ages > 150)) {
    stop("You have negative ages, or you're too old.")
  }
  mean(ages)
}

ages <- c(54, 45, 67, 151)

mean_age(ages)

```

# Chapter 3. Preparing your defences.

DRY - Do not repeat yourself. 
WET - Write everything twice (should be avoided). We enjoy typing. 

Comments - avoid comments you will never update. Be consistent.

Avoid using '.' (full stops) in variable or function names. They have a special meaning in R.

Don't mix and match assignment (e.g. <- or =)

Consistent spacing makes code far easier to read.
- use spaces around assignment arrows
- use spaces after a comma

The `lintr` package offers static code analysis for R. Similar to a spell check, but for formatting. Save your code (.R) and run lintr on it.

# Chapter 4. A Battle Plan

Consistency in file names: 
+ Dashes seem to be the best. Underscores are ok. Spaces are terrible.
Organizing a project:
+ Every project - has its own directory, has a sensible file structure.
+ R/, load.R, clean.R, function.R, analysis.R ... etc. 
+ Don't use absolute paths. 
+ Graphics and output - output/ and graphics/










