---
title: "Developing R Packages"
author: "Marcus Becker"
date: "December 11, 2019"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(here)

```

# Chapter 1: The R Package Structure

An R Packages can include:
+ Functions
+ Data
+ Documentation
+ Vignettes
+ Tests

What must be included?

+ R directory
+ man directory -> for documentation. 
+ NAMESPACE file
+ DESCRIPTION file

Packages for easily creating packages!

+ devtools
  - You can use devtools to create the basic structure of an R package by using the create() function.
+ roxygen2 


```{r}

library(devtools)

# Use the create() function to create a new R package called datasummary
create("C:/Users/mabec/Documents/R/datasummary")

dir("C:/Users/mabec/Documents/R/datasummary")

```

Now, let's create a function for our package. This goes in the R folder. Note that you can't have subdirectories in this folder.

Other tips:
- Don't have everything in one script.
- Don't have a large number of small files.
- Group similar functions together.

```{r}

# Create numeric_summary() function
numeric_summary <- function(x, na.rm) {

  # Include an error if x is not numeric
  if(!is.numeric(x)){
    stop("Data must be numeric")
  }

  # Create data frame
  else(data.frame(min = min(x, na.rm = na.rm),
                  median = median(x, na.rm = na.rm),
                  sd = sd(x, na.rm = na.rm),
                  max = max(x, na.rm = na.rm)))
}

# Then we dump it into the R folder
dump("numeric_summary", file = "C:/Users/mabec/Documents/R/datasummary/R/numeric_summary.R")



```

The use_* functions:

Beyond the required structure, you can include a number of additional directories containing elements such as vignettes (user guides), data, and unit tests. 

```{r}

# Include data
weather <- data.frame(day = 1:7,
                      Temp = sample(10:25, 7),
                      Weather = sample(c("Snow", "Rain", "Sun"), 7, replace = TRUE))

use_data(weather, pkg = "datasummary", overwrite = TRUE)

# Include vignettes
use_vignette("my_first_vignette", pkg = "simutils") # Rmd template

```

# Chapter 2: Documentation

Introduction to roxygen2

The `roxygen` headers are included in the same script as the function code. You can use roxygen comments (#') to identify a line as part of the roxygen header. 

The first three lines of the header have special meaning, and don't require tags to identify them. (use roxygen comments tho, and have a blank line between them)
+ title
+ description
+ details

Use the @param tag to document the arguments of your function.

If you wish to use functions from other packages, you can use the @import in the roxygen header (or @importFrom to import a single function)

How do we export functions?

Exported functions:
- Visible to the end user
- key to package functionality

Non-exported functions:
- not visible to the end user
- utility functions

Use the @export tag in the roxygen header to indicate that the function should be exported.

Examples are the way that your end users will learn how to work with your code.













