---
title: "Hierarchical and Mixed Effects Models in R"
author: "Marcus Becker"
date: "February 15, 2020"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)

```

# Chapter 1: Overview and Introduction to Hierarchical and Mixed Models

- lme4 package

```{r}

library(lme4)
library(lmerTest)
library(broom.mixed)

studentData <- read_csv("https://assets.datacamp.com/production/repositories/1803/datasets/975fe2b0190804d854a5da90083364629fb6af2e/classroom.csv")

# Visualize the data first
ggplot(data = studentData, aes(x = housepov, y = mathgain)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)

# Fit a linear model
summary(lm(formula = mathgain ~ housepov, data = studentData)) # Not predictive. We haven't accounted for the hierarchical nature of the data.

# Class
classData <- studentData %>%
  group_by(classid) %>%
  summarise(mathgain = mean(mathgain),
            housepov = mean(housepov))

ggplot(data = classData, aes(x = housepov, y = mathgain)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)

# School
schoolData <- studentData %>%
  group_by(schoolid) %>%
  summarise(mathgain = mean(mathgain),
            housepov = mean(housepov))

ggplot(data = schoolData, aes(x = housepov, y = mathgain)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)

```

Multiple regression in R tips:

+ lm(y ~ x -1) - the -1 estimates an intercept for each x (group), rather than relative to the first group.
+ numeric vs factors - R automatically assumes a numeric predictor is a slope
+ scaling parameters and slopes
+ shortcut for interaction is x1*x2

Without other coefficients, a single intercept is the global mean of the data.
Multiple intercepts allow you to estimate the mean for each group as long as other coefficients are not estimated (when the groups are discrete...).

What about continuous predictor variables? -> slopes.

```{r}

# Mixed effects model (i.e. has both fixed and random effects)
me_mod <- lmer(formula = mathgain ~ sex + mathprep + mathknow + (1|classid) + (1|schoolid),
               data = studentData,
               na.action = "na.omit",
               REML = TRUE)

summary(me_mod)

# Extract out the coefficients
modelOutPlot <- broom.mixed::tidy(me_mod, conf.int = TRUE) # note: have to use broom.extra for me models

modelOutPlot <- modelOutPlot[modelOutPlot$effect == "fixed" &
                             modelOutPlot$term != "(Intercept)", ]

# Plot the coefficients of interest
ggplot(data = modelOutPlot, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  theme_minimal() +
  geom_hline(yintercept = 0.0, color = "red", size = 2) +
  geom_point() +
  geom_linerange() +
  coord_flip()

# ^ Really like this plot!

```

Chapter 2: Linear Mixed-Effect Models

Random-effect syntax (for using lme4):

- (1 | group): random intercept with a fixed mean
- (1 | g1/g2): intercepts vary among g1 and g2 within g2
- (1 | g1) + (1 | g2): random intercepts for two variables
- x + (x | g): correlated random slope and intercept
- x + (x || g): uncorrelated random slope and intercept

```{r}

# Using birth data
countyBirthsData <- read_csv("https://assets.datacamp.com/production/repositories/1803/datasets/eb95cb6973afa56c38ba53cfd8058c72f768322f/countyBirthsDataUse.csv")

# Counties exist within states, and perhaps states contribute to variability. Hence, the need for random effects.

# To start, we build a hierarchical model with a global intercept (fixed-effect) and random-effect for state.
mod1 <- lmer(formula = BirthRate ~ (1 | State),
             data = countyBirthsData)

summary(mod1)
# plot the residuals
plot(mod1)

countyBirthsData$predictState <- predict(mod1, countyBirthsData)

ggplot(data = countyBirthsData) +
  theme_minimal() +
  geom_point(aes(x = TotalPopulation, y = BirthRate)) +
  geom_point(aes(x = TotalPopulation, y = predictState),
             color = "blue", alpha = 0.5)

# Random-effects intercepts estimated for each state. This allowed us to account for each state having its own intercept.

# Let's include a fixed effect: average age of mother.

mod2 <- lmer(formula = BirthRate ~ AverageAgeofMother + (1 | State), 
             data = countyBirthsData)

summary(mod2) 

# Now we want a random-effects slope for each state. A random-effect slope may be estimated for each group using the (slope | group) syntax. 

# Adding total population of each state as a random effect (it's numeric, hence slope not intercept)

countyBirthsData <- countyBirthsData %>%
  mutate(LogTotalPop = log10(TotalPopulation))

mod3 <- lmer(formula = BirthRate ~ AverageAgeofMother + (LogTotalPop | State),
             data = countyBirthsData)

# Uncorrelated random effects model (use the || syntax):
mod4 <- lmer(formula = BirthRate ~ AverageAgeofMother + (LogTotalPop || State),
             data = countyBirthsData)

summary(mod4) # Not good. Need to the more complex model (mod3)

# A predictor can be both a fixed-effect and a random-effect (e.g., mothers age)
mod5 <- lmer(formula = BirthRate ~ AverageAgeofMother + (AverageAgeofMother | State),
             data = countyBirthsData)

summary(mod5) # This is helpful for prediction, as you can correct the effect of age of mother by state (diff slope estimates by state)



```

REML - restricted maximum likelihood method

```{r}

# Extracting coefficients

# Fixed effects:
fixef(mod5)
# Random effects:
ranef(mod5)
# Confidence intervals for fe:
confint(mod5)


tidy(mod5, conf.int = TRUE)

```

Using Maryland Crime data

```{r}

MDcrime <- read_csv("https://assets.datacamp.com/production/repositories/1803/datasets/e5e076efd3c3b7665a3180da9f95aaaf671f6a61/MDcrime.csv")

ggplot(data = MDcrime, aes(x = Year, y = Crime, group = County)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  ylab("Major crimes reported per country across time.")

# Looks like we will require a random-effect intercept, and likely a random-effect slope. 

# Try a linear model:
mod1 <- lm(formula = Crime ~ Year, data = MDcrime)
summary(mod1) # Year(2) not significant. Note: need to use Year2 variable.

# Fit the model with Year(2) as both a fixed and random-effect:
mod2 <- lmer(formula = Crime ~ Year2 + ( 1 + Year2 | County), data = MDcrime)
summary(mod2)

# ANOVA - Analysis of Variance
# Build a null model with only County as a random-effect
null_model <- lmer(Crime ~ (1 | County), data = MDcrime)
# Build alternative model, with Year2 as a fixed and random slope and County as a random effect
alt_model <- lmer(Crime ~ Year2 + (1 + Year2 | County), data = MDcrime)

# Compare models - look at Chi square test.
anova(null_model, alt_model)

```

# Chapter 3: Generalized Linear Mixed-Effects Models

Topics covered in this chapter (for future reference):

+ Logisitic regression with glm() - family = "binomial"
+ Poisson regression with glm() - family = "poisson" -> for count data.
+ Plotting logistic regression results with ggplot - stat_smooth(method = "glm", method.args = list(family = "binomial"))
+ Using glmer() to estimate glms with mixed effects.
+ Handling different data inputs (e.g. matrix with cbind)
+ Calculating odds-ratios with exp(fixef()) and exp(confint()). Can also be done via tidy()
+ Using ggplot to visualize random-effects poisson models 

# Chapter 4: Repeated Measures

Paired t-test -> special case of a t-test

t.test(paired = TRUE) -> does not assume equal variance of both groups. 

Repeated measures ANOVA -> tests if means are constant across time.
^ simply a special type of mixed effects model.

```{r}

set.seed(1234)

n_ind <- 10
before <- rnorm(n = n_ind, mean = 0, sd = 0.5)
after <- before + rnorm(n = n_ind, mean = 4.5, sd = 5)

t.test(before, after, paired = F)
t.test(before, after, paired = T) # Paired is more powerful. 

dat <- data.frame(y = c(before, after),
                  trial = rep(c("before", "after"), each = n_ind),
                  ind = rep(letters[1:n_ind], times = 2))

# Now do it with lmer:
mod <- lmer(y ~ trial + (1|ind), data = dat)
summary(mod) # Woot. Pretty much the same (same t value, p value)



```

```{r}

htcrime <- read_csv("https://assets.datacamp.com/production/repositories/1803/datasets/45e88fe1bc8d1d76d140e69cb873da9eddb7008e/hateNY.csv")

# Is the number of hate crimes changing over time in NY counties? 
# 1) Is the state-wide number of hate crimes changing?
# 2) Are the number of hate crimes changing differently in each county?

# Step 1. Visualize

ggplot(data = htcrime, mapping = aes(x = Year, y = TotalIncidents, group = County)) +
  geom_line() +
  geom_smooth(method = "glm", method.args = c("poisson"), se = FALSE)

# We can see that different counties have different trends. Therefore it would be appropriate to use different random-effect slopes for each group (county).



```









